A Statistical Dynamical Model to Predict Extreme Events and Anomalous Features in Shallow Water Waves with Abrupt Depth Change

### 2018-11-27 paper1_tKdV/tKdV_theory.tex
NM

The Korteweg-de Vries (KdV) equation is a one-dimensional, deterministic model capable of describing {\color{blue} (weak?)} surface wave turbulence. More specifically, KdV is leading-order approximation for surface waves governed by a balance of nonlinear and dispersive effects, valid in an appropriate far-field limit \cite{johnson1997modern}. Moreover, KdV has been adapted to describe waves propagating over variable depth  \cite{johnson1997modern}. 
Here, we consider the variable-depth KdV equation truncated at wavenumber $\Lambda$ (with $J=2\Lambda+1$ grid points) in order to generate weakly turbulent dynamics. The surface displacement is described by the state variable $u_{\Lambda}^{\pm}\left(x,t\right)$ with superscript `$-$' for the incoming waves and `$+$' for the outgoing waves.
The Galerkin truncated variable $u_{\Lambda}=\sum_{1\leq\left|k\right|\leq\Lambda}\hat{u}_{k}\left(t\right)e^{ikx}$ is normalized with zero mean $\hat{u}_{0}=0$ and unit energy $2\pi\sum_{k=1}^{\Lambda}\left|\hat{u}_{k}\right|^{2}=1$, which are conserved quantities. Here, $u_{\Lambda}\equiv\mathcal{P}_{\Lambda}u$ denotes the subspace projection. The evolution of  $u_{\Lambda}^{\pm}$ is governed by the truncated KdV equation with depth change $D_{\pm}$
\begin{equation}
\frac{\partial u_{\Lambda}^{\pm}}{\partial t}+\frac{D_{\pm}^{-3/2}}{2}E_{0}^{1/2}L_{0}^{-3/2}\frac{\partial}{\partial x}\mathcal{P}_{\Lambda}\left(u_{\Lambda}^{\pm}\right)^{2}+D_{\pm}^{1/2}L_{0}^{-3}\frac{\partial^{3}u_{\Lambda}^{\pm}}{\partial x^{3}}=0. \label{eq:tKdV_nondim}
\end{equation}
Equation [\ref{eq:tKdV_nondim}] is non-dimensionalized on the periodic
domain $x\in\left[-\pi,\pi\right]$. The depth is assumed to be unit $D_{-}=1$ before the ADC and $D_{+}<1$ after the ADC. 
The conserved Hamiltonian can be decomposed as
\begin{align*}
\mathcal{H}_{\Lambda}^{\pm} & =  D_{\pm}^{-3/2}E_{0}^{1/2}L_{0}^{-3/2}H_{3}\left(u_{\Lambda}^{\pm}\right)-D_{\pm}^{1/2}L_{0}^{-3}H_{2}\left(u_{\Lambda}^{\pm}\right),\\
& H_{3}\left(u\right) =  \frac{1}{6}\int_{-\pi}^{\pi}u^{3}dx,\;H_{2}\left(u\right)=\frac{1}{2}\int_{-\pi}^{\pi}\left(\frac{\partial u}{\partial x}\right)^{2}dx.
\end{align*}
where we refer to $H_3$ as the cubic term and $H_2$ the quadratic term.
We introduce parameters $\left(E_{0},L_{0},\Lambda\right)$
based on the following assumptions:
\begin{itemize}
\item The wavenumber truncation $\Lambda$ is fixed at a moderate value
for generating weakly turbulent dynamics;
\item The state variable $u_{\Lambda}^{\pm}$ is normalized with zero mean
and unit energy, $\mathcal{M}\left(u_{\Lambda}\right)=0,\mathcal{E}\left(u_{\Lambda}\right)=1$,
which are conserved during evolution. Meanwhile, $E_{0}$ characterizes the total
energy injected into the system based on the driving amplitude $\Delta \theta$;
\item The length scale of the system $L_{0}$ is chosen so that the resolved scale $\Delta x=2\pi L_{0}/J$ is comparable to the the characteristic wave length $\lambda_{c}$ from the experiments.
\end{itemize}
% ORIGINAL
%The intuition for the distinct model dynamics comes from the balance
%between the cubic and quadratic terms in the Hamiltonian $\mathcal{H}_{\Lambda}^{\pm}$.
%After the depth change, $D_{+}<1$, more weight is added in the cubic
%term, $H_{3}$, for stronger nonlinearity and weaker dispersion for
%the third-order derivative term reflected by the smaller coefficient
%for $H_{2}$ in the Hamiltonian. 
Some intuition for how \eqref{eq:tKdV_nondim} produces different dynamics on either side of the ADC can be gained by considering the relative contributions of the cubic and quadratic terms to the Hamiltonian $\mathcal{H}_{\Lambda}^{\pm}$. The depth change, $D_{+}<1$, increases the weight of $H_3$ and decreases that of $H_2$, thus promoting the effects of nonlinearity and reducing dispersion.
Since $\frac{\partial u}{\partial x}$ is the slope of the wave height, $H_{2}\left(u\right)$ measures the
wave slope energy.

>>>>

The Korteweg-de Vries (KdV) equation is a one-dimensional, deterministic model capable of describing {\color{blue} (weak?)} surface wave turbulence. More specifically, KdV is leading-order approximation for surface waves governed by a balance of nonlinear and dispersive effects, valid in an appropriate far-field limit \cite{johnson1997modern}. Moreover, KdV has been adapted to describe waves propagating over variable depth  \cite{johnson1997modern}. 
Here, we consider the variable-depth KdV equation truncated at wavenumber $\Lambda$ (with $J=2\Lambda+1$ grid points) in order to generate weakly turbulent dynamics. The surface displacement is described by the state variable $u_{\Lambda}^{\pm}\left(x,t\right)$ with superscript `$-$' for the incoming waves and `$+$' for the outgoing waves.
The Galerkin truncated variable $u_{\Lambda}=\sum_{1\leq\left|k\right|\leq\Lambda}\hat{u}_{k}\left(t\right)e^{ikx}$ is normalized with zero mean $\hat{u}_{0}=0$ and unit energy $2\pi\sum_{k=1}^{\Lambda}\left|\hat{u}_{k}\right|^{2}=1$, which are conserved quantities. Here, $u_{\Lambda}\equiv\mathcal{P}_{\Lambda}u$ denotes the subspace projection. The evolution of  $u_{\Lambda}^{\pm}$ is governed by the truncated KdV equation with depth change $D_{\pm}$
\begin{equation}
\frac{\partial u_{\Lambda}^{\pm}}{\partial t}+\frac{D_{\pm}^{-3/2}}{2}E_{0}^{1/2}L_{0}^{-3/2}\frac{\partial}{\partial x}\mathcal{P}_{\Lambda}\left(u_{\Lambda}^{\pm}\right)^{2}+D_{\pm}^{1/2}L_{0}^{-3}\frac{\partial^{3}u_{\Lambda}^{\pm}}{\partial x^{3}}=0. \label{eq:tKdV_nondim}
\end{equation}
Equation [\ref{eq:tKdV_nondim}] is non-dimensionalized on the periodic
domain $x\in\left[-\pi,\pi\right]$. The depth is assumed to be unit $D_{-}=1$ before the ADC and $D_{+}<1$ after the ADC. 
The conserved Hamiltonian can be decomposed as
\begin{align*}
\mathcal{H}_{\Lambda}^{\pm} & =  D_{\pm}^{-3/2}E_{0}^{1/2}L_{0}^{-3/2}H_{3}\left(u_{\Lambda}^{\pm}\right)-D_{\pm}^{1/2}L_{0}^{-3}H_{2}\left(u_{\Lambda}^{\pm}\right),\\
& H_{3}\left(u\right) =  \frac{1}{6}\int_{-\pi}^{\pi}u^{3}dx,\;H_{2}\left(u\right)=\frac{1}{2}\int_{-\pi}^{\pi}\left(\frac{\partial u}{\partial x}\right)^{2}dx.
\end{align*}
where we refer to $H_3$ as the cubic term and $H_2$ the quadratic term.
We introduce parameters $\left(E_{0},L_{0},\Lambda\right)$
based on the following assumptions:
\begin{itemize}
\item The wavenumber truncation $\Lambda$ is fixed at a moderate value
for generating weakly turbulent dynamics.
\item The state variable $u_{\Lambda}^{\pm}$ is normalized with zero mean
and unit energy, $\mathcal{M}\left(u_{\Lambda}\right)=0,\mathcal{E}\left(u_{\Lambda}\right)=1$,
which are conserved during evolution. Meanwhile, $E_{0}$ characterizes the total
energy injected into the system based on the driving amplitude $\Delta \theta$.
\item The length scale of the system $L_{0}$ is chosen so that the resolved scale $\Delta x=2\pi L_{0}/J$ is comparable to the the characteristic wave length $\lambda_{c}$ from the experiments.
\end{itemize}
% ORIGINAL
%The intuition for the distinct model dynamics comes from the balance
%between the cubic and quadratic terms in the Hamiltonian $\mathcal{H}_{\Lambda}^{\pm}$.
%After the depth change, $D_{+}<1$, more weight is added in the cubic
%term, $H_{3}$, for stronger nonlinearity and weaker dispersion for
%the third-order derivative term reflected by the smaller coefficient
%for $H_{2}$ in the Hamiltonian. 
Some intuition for how \eqref{eq:tKdV_nondim} produces different dynamics on either side of the ADC can be gained by considering the relative contributions of $H_3$ and $H_2$ in the Hamiltonian $\mathcal{H}_{\Lambda}^{\pm}$. The depth change, $D_{+}<1$, increases the weight of $H_3$ and decreases that of $H_2$, thereby promoting the effects of nonlinearity over dispersion and creating conditions favorable for extreme events. Since $\frac{\partial u}{\partial x}$ is the slope of the wave height, $H_{2}\left(u\right)$ measures the
wave slope energy.
